---
title: "Explore Soil Core Data"
output: html_document
---
```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r setup}
library(tidyverse)
library(sf)
library(tmap)
library(DT)
library(gstat)
library(raster)
library(maptools)
#map_cookeast_idw(sf.bd1998, cookeast, AvgBD, seq(1.1,1.7,by=0.1), cols <- blues9[3:9])
```

```{r idwMap}
map_cookeast_idw <- function(simpleFeatureObj, 
                             cookEastBoundary, 
                             factor,
                             cutSeq,
                             colSeq) {
  # TODO: add params for hardcoded values
  # Convert sf to sp
  sp <- as(simpleFeatureObj, "Spatial")
  
  # Create grid for IDW
  e <- extent(sp)
  grid <- raster(e)
  proj4string(grid) <- proj4string(sp)
  res(grid) <- c(0.00004, 0.00004)
  grid <- as(grid, "SpatialPointsDataFrame")
  
  # Passing variable names is tricky: https://stackoverflow.com/a/34079727
  var <- eval(substitute(factor), simpleFeatureObj)
  var.name <- substitute(var)
  
  # Generate IDW 
  idw <- idw(formula = var ~ 1, locations = sp, newdata = grid, idp = 2)
  
  # Clip and map
  idw.sf.clip <- st_intersection(st_as_sf(idw), cookEastBoundary)
  idw.clip <- as(idw.sf.clip, "Spatial")
  
  #cuts <- 0.2*(1.2:7)
  
  if(is.na(cutSeq) | is.na(colSeq)) {
    spplot(idw.clip, 
         "var1.pred", 
         sp.layout = list("sp.points", sp, pch = 16, cex = 0.5, col = "black"))
  } else {
    spplot(idw.clip, 
         "var1.pred", 
         cuts = cutSeq, 
         col.regions = colSeq,
         sp.layout = list("sp.points", sp, pch = 16, cex = 0.5, col = "black"))
  }
  
  #ggplot() + geom_sf(data = idw.sf.clip, aes(col = var1.pred))
  #par(mfrow=c(1,2),oma=c(2,2,2,2))
  #cols = bpy.colors(30)
  #zmin = 0
  #zmax = 100
  #image(idw.clip, zlim=c(zmin,zmax),col=cols)
  #box()
  #mtext(side=3,text="zinc")
  #image.plot(zlim=c(zmin,zmax),nlevel=5,legend.only=TRUE,horizonal=F,col=cols)
}
```

## Load data

```{r loadData}
df <- read_csv("../output/soilCore1998To2015ShallowDeepMergedByHorizon_20201217.csv")
sf <- st_as_sf(df, coords = c("Longitude", "Latitude"), crs = 4326)
cookeast <- st_read("../input/CookEastBoundary_20180131.geojson")
```

Data fields

```{r headers}
df %>% names()
```


## Map sample locations

Demo of 2015 data
```{r}
tmap_mode("view")
sf %>% 
  filter(Year == 2015) %>% 
  group_by(ID2) %>% 
  summarize(avgBD = mean(BulkDensity), totalTNStock = sum(TNStock), totalTocStock = sum(TocStock)) %>% 
  tm_shape() +
  tm_symbols(size = 0.5) +
  tm_basemap("https://server.arcgisonline.com/ArcGIS/rest/services/World_Topo_Map/MapServer/tile/{z}/{y}/{x}")
```
Compare samples across years

```{r ggplotMap }
sf %>% ggplot() + 
  geom_sf(aes(color = as.factor(Year))) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_grid(. ~ Year)
```

# Bulk Density (g/cm^3)

```{r bulkDensityWideFormat}
#bulkDensity <- df %>% 
#  dplyr::select(ID2, Year, TopDepth, BottomDepth, BulkDensity) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  mutate(key = row_number()) %>% 
#  spread(Year, BulkDensity) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  fill(-ID2, TopDepth, BottomDepth) %>% 
#  mutate(maxKey = max(key)) %>% 
#  filter(key == maxKey) %>% 
#  dplyr::select(-key, -maxKey) %>% 
#  rename(BulkDensity1998 = 4, BulkDensity2005 = 5, BulkDensity2015 = 6)
```

Display table

```{r bulkDensityTable}
#bulkDensity %>% datatable(rownames = F)
```

Map avg bulk density

```{r bulkDensityMap, out.width="100%"}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, BulkDensity) %>%  
  group_by(Year, ID2) %>% 
  summarize(AvgBD = mean(BulkDensity)) %>% 
  ggplot() + 
    geom_sf(aes(color = AvgBD)) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    facet_grid(Year ~ .)
```

### IDW Bulk Density 1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, BulkDensity) %>%  
  group_by(Year, ID2) %>% 
  summarize(AvgBD = mean(BulkDensity)) %>% 
  filter(!is.na(AvgBD)) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, AvgBD, seq(1.1,1.7,by=0.1), cols <- blues9[3:9])
```

### IDW Bulk Density 2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  dplyr::select(ID2, TopDepth, BottomDepth, BulkDensity) %>%  
  group_by(ID2) %>% 
  summarize(AvgBD = mean(BulkDensity)) %>% 
  filter(!is.na(AvgBD)) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, AvgBD, seq(1.1,1.7,by=0.1), cols <- blues9[3:9])
```

### IDW Bulk Density 2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  dplyr::select(ID2, TopDepth, BottomDepth, BulkDensity) %>%  
  group_by(ID2) %>% 
  summarize(AvgBD = mean(BulkDensity)) %>% 
  filter(!is.na(AvgBD)) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, AvgBD, seq(1.1,1.7,by=0.1), cols <- blues9[3:9])
```

# pH

```{r pHWideFormat}
#pH <- df %>% 
#  dplyr::select(ID2, Year, TopDepth, BottomDepth, pH) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  mutate(key = row_number()) %>% 
#  spread(Year, pH) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  fill(-ID2, TopDepth, BottomDepth) %>% 
#  mutate(maxKey = max(key)) %>% 
#  filter(key == maxKey) %>% 
#  dplyr::select(-key, -maxKey) %>% 
#  rename(pH1998 = 4, pH2008 = 5, pH2015 = 6)
```

Display table

```{r pHTable}
#pH %>% datatable(rownames = F)
```

Map pH, year by bottom depth (0-10 cm, 10-20 cm, 20-30 cm)

```{r pHMap, out.width="100%"}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, pH) %>%  
  filter(BottomDepth < 31) %>% 
  na.exclude() %>% 
  ggplot() + 
    geom_sf(aes(color = pH)) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    facet_grid(Year ~ BottomDepth)
```

### IDW pH 1998 @ 20-30 cm depth increment

```{r}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, pH) %>% 
  filter(Year == 1998) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, pH, seq(2,8,by=1), cols <- blues9[3:9])
```

### IDW pH 2008 @ 20-30 cm depth increment

```{r}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, pH) %>% 
  filter(Year == 2008) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, pH, seq(2,8,by=1), cols <- blues9[3:9])
```

### IDW pH 2015 @ 20-30 cm depth increment

```{r}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, pH) %>% 
  filter(Year == 2015) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, pH, seq(2,8,by=1), cols <- blues9[3:9])
```

# Total organic carbon stock (Mg/ha)

```{r}
#tocStock <- df %>% 
#  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  mutate(key = row_number()) %>% 
#  spread(Year, TocStock) %>% 
#  dplyr::select(-key) %>% 
#  group_by(ID2, TopDepth, BottomDepth) %>% 
#  fill(-ID2, TopDepth, BottomDepth) %>% 
#  rename(TocStock1998 = 4, TocStock2005 = 5, TocStock2015 = 6) %>% 
#  filter(!is.na(TocStock2015))
```

Display table

```{r}
#tocStock %>% datatable(rownames = F)
```

Map profile TOC Stock

```{r, out.width="100%"}
sf %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>% 
  ggplot() + 
    geom_sf(aes(color = ProfileTocStock)) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    facet_grid(Year ~ .)
```

### IDW Profile TOC Stock 1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(80,260,by=30), cols <- blues9[3:9])
```

### IDW Profile TOC Stock 2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>% 
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(80,260,by=30), cols <- blues9[3:9])
```

### IDW Profile TOC Stock 2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(80,260,by=30), cols <- blues9[3:9])
```

### IDW 0 - 10 cm TOC Stocks

1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  filter(TopDepth == 0 & BottomDepth == 10) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(12.9,51.9,by=6.5), cols <- blues9[3:9])
```

2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  filter(TopDepth == 0 & BottomDepth == 10) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(12.9,51.9,by=6.5), cols <- blues9[3:9])
```

2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  filter(TopDepth == 0 & BottomDepth == 10) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(12.9,51.9,by=6.5), cols <- blues9[3:9])
```

### IDW 10 cm - 20 cm TOC Stocks

1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  filter(TopDepth == 10 & BottomDepth == 20) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(2.3,36,by=5.6), cols <- blues9[3:9])
```

2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  filter(TopDepth == 10 & BottomDepth == 20) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(2.3,36,by=5.6), cols <- blues9[3:9])
```

2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  filter(TopDepth == 10 & BottomDepth == 20) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(2.3,36,by=5.6), cols <- blues9[3:9])
```

### IDW 20 cm - 30 cm TOC Stocks

1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(4.1,42.7,by=6.4), cols <- blues9[3:9])
```

2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(4.1,42.7,by=6.4), cols <- blues9[3:9])
```

2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  filter(TopDepth == 20 & BottomDepth == 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(4.1,42.7,by=6.4), cols <- blues9[3:9])
```

### IDW 30 cm - 153 cm TOC Stocks

1998

```{r}
sf %>% 
  filter(Year == 1998) %>% 
  filter(TopDepth > 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(17.4,199,by=30.3), cols <- blues9[3:9])
```

2008

```{r}
sf %>% 
  filter(Year == 2008) %>% 
  filter(TopDepth > 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(17.4,199,by=30.3), cols <- blues9[3:9])
```

2015

```{r}
sf %>% 
  filter(Year == 2015) %>% 
  filter(TopDepth > 30) %>% 
  dplyr::select(ID2, Year, TopDepth, BottomDepth, TocStock) %>%  
  group_by(Year, ID2) %>% 
  summarize(ProfileTocStock = sum(TocStock)) %>%
  na.exclude() %>% 
  map_cookeast_idw(cookeast, ProfileTocStock, seq(17.4,199,by=30.3), cols <- blues9[3:9])
```